---
title: "Group Project #1: Prosper"
author: "Group 3: Adam John Altepeter, Anyu Lei, Shubkarman Singh Sidhu, Andrew Tseng, Takumi Umemaki"
date: "2021/3/28"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Exploration and Cleansing

### Data Structure and Summary

First, we import the dataset and see its structure.

```{r}
# Import data and see the summary
listings <- read.csv("ProjectA_Listings2013.csv")
str(listings)
summary(listings)
```

### Data Transformation and Cleansing

Then, we convert applicable variables into factors to develop both linear and logistic regression models.

```{r}
# Convert applicable variables into factors
listings$loan_status <- as.factor(listings$loan_status)
listings$loan_status <- factor(listings$loan_status, labels = c("CURRENT", "CHARGEOFF", "DEFAULTED", "COMPLETED"))
listings$prosper_rating <- as.factor(listings$prosper_rating)
listings$scorex <- as.factor(listings$scorex)
listings$listing_category_id <- as.factor(listings$listing_category_id)
listings$income_range <- as.factor(listings$income_range)
listings$income_range <- factor(listings$income_range, labels = c("$0", "$1-24,999", "$25,000-49,999", "$50,000-74,999", "$75,000-99,999", "$100,000+", "Not employed"))
listings$employment_status_description <- as.factor(listings$employment_status_description)
listings$occupation <- as.factor(listings$occupation)
listings$lender_indicator <- as.factor(listings$lender_indicator)
```

We also delete unnecessary columns.

```{r}
# Delete repeated variables
listings$loan_status_description <- NULL
listings$income_range_description <- NULL

# Delete date variables
listings$loan_origination_date <- NULL
listings$first_recorded_credit_line <- NULL #Assuming this date does not affect interest rates

# Delete location variables
listings$borrower_state <- NULL #Assuming borrower states do not affect interest rates
listings$borrower_city <- NULL #Assuming borrower cities do not affect interest rates
```

Finally, we convert negative values.

```{r}
# Set negative values in months_employed to NA
listings$months_employed <- ifelse(listings$months_employed < 0, NA, listings$months_employed)

# See updated data summary
summary(listings)
```

***

## Linear Regression Model {.tabset}

### Model 1

> Include all factors as independent variables in the linear regression model (Starting Model)

```{r}
# Develop a linear regression model
model1 <- lm(borrower_rate ~ ., data = listings)
summary(model1)
```

### Model 2

> Remove all insignificant variables from Model 1

```{r}
# Remove insignificant variables
model2 <- lm(borrower_rate ~ number_of_days + principal_balance + loan_status + amount_funded + prosper_rating + listing_term + listing_monthly_payment + scorex + stated_monthly_income + employment_status_description + occupation + lender_indicator + monthly_debt + inquiries_last6_months + open_credit_lines + bankcard_utilization + delinquencies_over30_days, data = listings)
summary(model2)
```

### Model 3

>  Remove ex post (items determined after loan funding) variables

```{r}
model3 <- lm(borrower_rate ~ loan_status + amount_funded + prosper_rating + listing_term + listing_monthly_payment + scorex + stated_monthly_income + employment_status_description + occupation + lender_indicator + monthly_debt + inquiries_last6_months + open_credit_lines + bankcard_utilization + delinquencies_over30_days, data = listings)
summary(model3)
```

### Model 4

> Use stepwise regression (forward and backward selection) to determine the appropriate variables 

```{r}
model4 <- lm(borrower_rate ~ amount_funded + prosper_rating + listing_term + listing_monthly_payment + scorex + prosper_score + income_verifiable + dti_wprosper_loan + lender_indicator + monthly_debt + current_delinquencies + public_records_last12_months + inquiries_last6_months + amount_delinquent + current_credit_lines + open_credit_lines + bankcard_utilization + total_open_revolving_accounts + real_estate_balance + total_inquiries + satisfactory_accounts + was_delinquent_derog + delinquencies_over30_days + delinquencies_over90_days + is_homeowner, data = listings)
summary(model4)
```

### Model Selection

> Choice: **Model 4**

* We start from Model 1 by including all potential independent variables to get an sense about how these variables help to explain the variations in `borrower_rate`. Then, we remove all insignificant variables to fit the model better. We also try adding one variable each time to the model to see if these variables become significant.
* We choose **Model 4** as our final regression model which explains the variations in loan interest rates because:
  * All independent variables except `scorex` are **significant** at a minimum of 95% level.
  * The model yields the highest **adjusted R-squared**, suggesting the model fits the data well. 

### Model Interpretation

> ***Overall Fit***

* **Model 4** has the highest adjusted R-squared among all candidates, so we expect this model to best predict further loan interest rates (`borrower_rate`).
* The adjusted R-squared of **Model 4** is `r summary(model4)$adj.r.squared`, indicting **`r paste(round(summary(model4)$adj.r.squared*100, 2), "%", sep="")`** of the variation in the loan interest rates (`borrower_rate`) in the sample is explained by the variation in a combination of loan term and size, credit risk and borrower status variables.

> ***Regression Coefficients***

```{r echo=FALSE, include=FALSE}
# All coefficients of Linear Regression Model 4
model4$coefficients
```

##### A. Loan Size and Term

1. `amount_funded`
* `amount_funded` **decreases** `borrower_rate`: Every one unit ($1) increase in the funded amount of the listing is associated with an estimated decrease in the loan interest rate of `r paste(round(-coef(summary(model4))["amount_funded","Estimate"], 4), "%", sep="")`.

2. `listing_term`
* `listing_term` **increase** `borrower_rate`: Every one unit (a month) increase in the listing terms is associated with an estimated increase in the loan interest rate of `r paste(round(coef(summary(model4))["listing_term","Estimate"], 4), "%", sep="")`.

3. `listing_monthly_payment`
* `listing_monthly_payment` **increase** `borrower_rate`: Every one unit ($1) increase in the listing monthly payment is associated with an estimated increase in the loan interest rate of `r paste(round(coef(summary(model4))["listing_monthly_payment","Estimate"], 4), "%", sep="")`.

##### B. Credit Score Matrices

1. `prosper_rating`
* `borrower_rate` of `prosper_ratingAA` is **lower**  than `prosper_ratingA`: Controlling for all other variables, the expected loan interest rate is `r paste(round(-coef(summary(model4))["prosper_ratingAA","Estimate"], 4), "%", sep="")` lower for a prosper rating of AA than for a prosper rating of A.
* `borrower_rate` of `prosper_ratingB` is **higher** than `prosper_ratingA`: Controlling for all other variables, the expected loan interest rate is `r paste(round(coef(summary(model4))["prosper_ratingB","Estimate"], 4), "%", sep="")` higher for a prosper rating of B than for a prosper rating of A.
* `borrower_rate` of `prosper_ratingC` is **higher** than `prosper_ratingA`: Controlling for all other variables, the expected loan interest rate is `r paste(round(coef(summary(model4))["prosper_ratingC","Estimate"], 4), "%", sep="")` higher for a prosper rating of C than for a prosper rating of A.
* `borrower_rate` of `prosper_ratingD` is **higher** than `prosper_ratingA`: Controlling for all other variables, the expected loan interest rate is `r paste(round(coef(summary(model4))["prosper_ratingD","Estimate"], 4), "%", sep="")` higher for a prosper rating of D than for a prosper rating of A.
* `borrower_rate` of `prosper_ratingE` is **higher** than `prosper_ratingA`: Controlling for all other variables, the expected loan interest rate is `r paste(round(coef(summary(model4))["prosper_ratingE","Estimate"], 4), "%", sep="")` higher for a prosper rating of E than for a prosper rating of A.
* `borrower_rate` of `prosper_ratingHR` is **higher** than `prosper_ratingA`: Controlling for all other variables, the expected loan interest rate is `r paste(round(coef(summary(model4))["prosper_ratingHR","Estimate"], 4), "%", sep="")` higher for a prosper rating of HR than for a prosper rating of A.

2. `scorex`
* `borrower_rate` of `scorex600-619` is **higher**  than `scorex<600`: Controlling for all other variables, the expected loan interest rate is `r paste(round(coef(summary(model4))["scorex600-619","Estimate"], 4), "%", sep="")` higher for a scorex range of 600-619 than for a scorex of lower than 600.
* `borrower_rate` of `scorex620-639` is **higher**  than `scorex<600`: Controlling for all other variables, the expected loan interest rate is `r paste(round(coef(summary(model4))["scorex620-639","Estimate"], 4), "%", sep="")` higher for a scorex range of 620-639 than for a scorex of lower than 600.
* `borrower_rate` of `scorex640-649` is **higher**  than `scorex<600`: Controlling for all other variables, the expected loan interest rate is `r paste(round(coef(summary(model4))["scorex640-649","Estimate"], 4), "%", sep="")` higher for a scorex range of 640-649 than for a scorex of lower than 600.
* `borrower_rate` of `scorex650-664` is **higher**  than `scorex<600`: Controlling for all other variables, the expected loan interest rate is `r paste(round(coef(summary(model4))["scorex650-664","Estimate"], 4), "%", sep="")` higher for a scorex range of 650-664 than for a scorex of lower than 600.
* `borrower_rate` of `scorex665-689` is **higher**  than `scorex<600`: Controlling for all other variables, the expected loan interest rate is `r paste(round(coef(summary(model4))["scorex665-689","Estimate"], 4), "%", sep="")` higher for a scorex range of 665-689 than for a scorex of lower than 600.
* `borrower_rate` of `scorex690-701` is **higher**  than `scorex<600`: Controlling for all other variables, the expected loan interest rate is `r paste(round(coef(summary(model4))["scorex690-701","Estimate"], 4), "%", sep="")` higher for a scorex range of 690-701 than for a scorex of lower than 600.
* `borrower_rate` of `scorex702-723` is **higher**  than `scorex<600`: Controlling for all other variables, the expected loan interest rate is `r paste(round(coef(summary(model4))["scorex702-723","Estimate"], 4), "%", sep="")` higher for a scorex range of 702-723 than for a scorex of lower than 600.
* difference in `borrower_rate` between `scorex724-747`, `scorex748-777`, `scorex778+` and base group `scorex<600` is not significant statistically.

3. `prosper_score`
* `prosper_score` **decreases** `borrower_rate`: Every one unit (one score) increase in the funded amount of the listing is associated with an estimated decrease in the loan interest rate of `r paste(round(-coef(summary(model4))["prosper_score","Estimate"], 4), "%", sep="")`.

##### C. Other Credit Risk Matrices

1. `dti_wprosper_loan`
* `dti_wprosper_loan` **increase** `borrower_rate`: Every one unit increase in the debt to income ratio is associated with an estimated increase in the loan interest rate of `r paste(round(coef(summary(model4))["dti_wprosper_loan","Estimate"], 4), "%", sep="")`.

2. `lender_indicator`
* `borrower_rate` of `lender_indicator1` is **higher**  than `lender_indicator0`: Controlling for all other variables, the expected loan interest rate is `r paste(round(coef(summary(model4))["lender_indicator1","Estimate"], 4), "%", sep="")` higher for a borrower who holds both borrower and investor roles than for a borrower who holds borrower role only.

3. `monthly_debt`
* `monthly_debt` **increase** `borrower_rate`: Every one unit ($1) increase in the monthly debt is associated with an estimated increase in the loan interest rate of `r paste(round(coef(summary(model4))["monthly_debt","Estimate"], 4), "%", sep="")`.

4. `current_delinquencies`
* `current_delinquencies` **increase** `borrower_rate`: Every one unit increase in the current delinquencies is associated with an estimated increase in the loan interest rate of `r paste(round(coef(summary(model4))["current_delinquencies","Estimate"], 4), "%", sep="")`.

5. `amount_delinquent`
* `amount_delinquent` **increase** `borrower_rate`: Every one unit ($1) increase in the current delinquent amount is associated with an estimated increase in the loan interest rate of `r paste(round(coef(summary(model4))["amount_delinquent","Estimate"], 4), "%", sep="")`.

6. `was_delinquent_derog`
* `was_delinquent_derog` **increase** `borrower_rate`: Every one unit increase in the previous delinquencies is associated with an estimated increase in the loan interest rate of `r paste(round(coef(summary(model4))["was_delinquent_derog","Estimate"], 4), "%", sep="")`.

7. `delinquencies_over30_days`
* `delinquencies_over30_days` **increase** `borrower_rate`: Every one unit increase in the number of delinquencies over 30 days at the time the listing is associated with an estimated increase in the loan interest rate of `r paste(round(coef(summary(model4))["delinquencies_over30_days","Estimate"], 4), "%", sep="")`.

8. `delinquencies_over90_days`
* `delinquencies_over90_days` **decrease** `borrower_rate`: Every one unit increase in the number of delinquencies over 90 days at the time the listing is associated with an estimated decrease in the loan interest rate of `r paste(round(-coef(summary(model4))["delinquencies_over90_days","Estimate"], 4), "%", sep="")`.

9. `public_records_last12_months`
* `public_records_last12_months` **increase** `borrower_rate`: Every one unit increase in the number of public records in the last 12 months at the time of listing is associated with an estimated increase in the loan interest rate of `r paste(round(coef(summary(model4))["public_records_last12_months","Estimate"], 4), "%", sep="")`.

10. `inquiries_last6_months`
* `inquiries_last6_months` **decrease** `borrower_rate`: Every one unit increase in the number of inquiries made in the last 6 months at the time the listing is associated with an estimated decrease in the loan interest rate of `r paste(round(-coef(summary(model4))["inquiries_last6_months","Estimate"], 4), "%", sep="")`.

11. `total_inquiries`
* `total_inquiries` **increase** `borrower_rate`: Every one unit increase in the total number of inquiries made is associated with an estimated increase in the loan interest rate of `r paste(round(coef(summary(model4))["total_inquiries","Estimate"], 4), "%", sep="")`.

12. `current_credit_lines`
* `current_credit_lines` **increase** `borrower_rate`: Every one unit increase in the number of current credit lines is associated with an estimated increase in the loan interest rate of `r paste(round(coef(summary(model4))["current_credit_lines","Estimate"], 4), "%", sep="")`.

13. `open_credit_lines`
* `open_credit_lines` **decrease** `borrower_rate`: Every one unit increase in the number of open credit lines is associated with an estimated decrease in the loan interest rate of `r paste(round(-coef(summary(model4))["open_credit_lines","Estimate"], 4), "%", sep="")`.

14. `bankcard_utilization`
* `bankcard_utilization` **increase** `borrower_rate`: Every one unit (1%) increase in the utilization rate of bankcard is associated with an estimated increase in the loan interest rate of `r paste(round(coef(summary(model4))["bankcard_utilization","Estimate"]/100, 5), "%", sep="")`.

15. `total_open_revolving_accounts`
* `total_open_revolving_accounts` **increase** `borrower_rate`: Every one unit increase in the total number of open revolving accounts is associated with an estimated increase in the loan interest rate of `r paste(round(coef(summary(model4))["total_open_revolving_accounts","Estimate"], 4), "%", sep="")`.

16. `satisfactory_accounts`
* `satisfactory_accounts` **decrease** `borrower_rate`: Every one unit increase in the number of satisfactory accounts is associated with an estimated decrease in the loan interest rate of `r paste(round(-coef(summary(model4))["satisfactory_accounts","Estimate"], 4), "%", sep="")`.

17. `real_estate_balance`
* `real_estate_balance` **increase** `borrower_rate`: Every one unit ($) increase in the real estate balance is associated with an estimated increase in the loan interest rate of `r paste(round(coef(summary(model4))["real_estate_balance","Estimate"], 4), "%", sep="")`.

18. `is_homeowner`
* `borrower_rate` of `is_homeownerTRUE` is **lower**  than `is_homeownerFALSE`: Controlling for all other variables, the expected loan interest rate is `r paste(round(-coef(summary(model4))["is_homeownerTRUE","Estimate"], 4), "%", sep="")` lower for a borrower who owns a home than for a borrower who does not.

##### D. Borrower Data

1. `income_verifiable`
* `borrower_rate` of `income_verifiableTRUE` is **higher**  than `income_verifiableFALSE`: Controlling for all other variables, the expected loan interest rate is `r paste(round(coef(summary(model4))["income_verifiableTRUE","Estimate"], 4), "%", sep="")` higher for a borrower whose income is verifiable than for a borrower whose income is not.

***

## Logistic Regression Model {.tabset}

#### Modify dataset for Logistic Regression

```{r}
# Set up new data frame
listings2 <- listings
listings2$loan_status <- ifelse(listings2$loan_status == "DEFAULTED", 1,0)
```

### Model 1

> Include all factors as independent variables in the logistic regression model (Starting Model)

```{r}
logisticmodel1 <- glm(loan_status~., data = listings2, family = "binomial")
summary(logisticmodel1)
```

### Model 2

> Remove all insignificant variables from Model 1

```{r}
logisticmodel2 = glm(loan_status ~ number_of_days + principal_balance + prosper_rating + borrower_rate + listing_term + bankcard_utilization + total_inquiries, data = listings2, family = "binomial")
summary(logisticmodel2)
```

### Model 3

> Use stepwise regression (forward and backward selection) to determine the appropriate variables

```{r}
logisticmodel3 = glm(loan_status ~ number_of_days + principal_balance + amount_funded + borrower_rate + listing_term + prosper_score + total_open_revolving_accounts + real_estate_balance + revolving_balance + total_inquiries + was_delinquent_derog, data = listings2, family = "binomial")
summary(logisticmodel3)
```

### Model 4

> Add `stated_monthly_income` to Model 3. There's a warning message of ***fitted probabilities numerically 0 or 1 occurred***. Since `stated_monthly_income` is positively skewed and contains lots of "0" values, we use log1p to transform the income data.

```{r}
# Set up new data frame and transform stated_monthly_income using log1p.
listings3 <- listings2
listings3$stated_monthly_income <- log1p(listings3$stated_monthly_income)

# Logistic regression model 4
logisticmodel4 = glm(loan_status ~ number_of_days + principal_balance + amount_funded + borrower_rate + listing_term + prosper_score + stated_monthly_income + total_open_revolving_accounts + real_estate_balance + total_inquiries + was_delinquent_derog, data = listings3, family = "binomial")
summary(logisticmodel4)
```

### Model Selection

> Choice: **Model 3**

* We start from Model 1 by including all potential independent variables to get an sense about how these variables help to explain the possibility of default. Then, we remove all insignificant variables to fit the model better. We also try adding one variable each time to the model to see if these variables become significant.
* We choose **Model 3** as final regression model which explains the variations in default rate because:
  * All independent variables are **significant** at a minimum of 95% level.
  * Using the model selection criteria from the summary table it appears that Model 2, Model 3, and Model 4 are very close. The AIC, Pseudo R-Squared, and Prediction Accuracy for the models are all within 1% of each other. As such, this does not show one model as superior to the others.
  * Specifically, **Model 3** yields relatively lower **AIC** and higher **McFadden's Pseudo R-squared**, suggesting **Model 3** fits the data better among the three models (not considering Model 1). 
  * Taking the predictability of fitted model into consideration, when we set the *cutoff point at 0.1*, **Model 3** has the highest accuracy in classifying default or not.
  * Looking at the Confusion matrices, the number of summed "False positives" and "False negatives," is lowest in model 2 (Model 2 = 2,293, Model 3 = 2,263, and Model 4 = 2,281). That said, the scenario that is most damaging is where you predict no default (a 0) and a default happens (a 1). The model that does the best in reducing this scenario is **Model 3**. As such, while there are very small differences in the models.
  * We set *cutoff = 0.1* because for all three models, the maximum value of predicted possibility of default is around 30%. Setting a higher cutoff point (e.g., 0.2) will lead most actual default borrowers being classified/predicted as not default.
  * *See below the comparisons of Model 2, 3 and 4.*

#### 1. Comparison: AIC

```{r}
AIC_02 = logisticmodel2$aic
AIC_03 = logisticmodel3$aic
AIC_04 = logisticmodel4$aic
```

#### 2. Comparison: McFadden's Pseudo R-squared

```{r}
pR2_02 = round(1 - logisticmodel2$deviance / logisticmodel2$null.deviance, 4)
pR2_03 = round(1 - logisticmodel3$deviance / logisticmodel3$null.deviance, 4)
pR2_04 = round(1 - logisticmodel4$deviance / logisticmodel4$null.deviance, 4)
```

#### 3. Comparison: Confusion Matrix and Classification Accuracy

```{r}
# Define a function to calculate error rate
calc_class_err = function(actual, predicted) {mean(actual != predicted)}

# Model 2: Classiifications using cutoff = 0.1, classification accuracy and error rate
logisticmodel2_pred = ifelse(predict(logisticmodel2, type = "response") > 0.1, 1, 0)
error_02 <- calc_class_err(actual = listings2$loan_status, predicted = logisticmodel2_pred)
accuracy_02 = round(1 - error_02, 5)

# Model 3: Classiifications using cutoff = 0.1, classification accuracy and error rate
logisticmodel3_pred = ifelse(predict(logisticmodel3, type = "response") > 0.1, 1, 0)
error_03 <- calc_class_err(actual = listings2$loan_status, predicted = logisticmodel3_pred)
accuracy_03 = round(1 - error_03, 5)

# Model 4: Classiifications using cutoff = 0.1, classification accuracy and error rate
logisticmodel4_pred = ifelse(predict(logisticmodel4, type = "response") > 0.1, 1, 0)
error_04 <- calc_class_err(actual = listings3$loan_status, predicted = logisticmodel4_pred)
accuracy_04 = round(1 - error_04, 5)
```

```{r}
# Model 2: Confusion Matrix
logisticmodel2_conf_table = table(predicted = logisticmodel2_pred, actual = listings2$loan_status)
logisticmodel2_conf_table

# Model 3: Confusion Matrix
logisticmodel3_conf_table = table(predicted = logisticmodel3_pred, actual = listings2$loan_status)
logisticmodel3_conf_table

# Model 4: Confusion Matrix
logisticmodel4_conf_table = table(predicted = logisticmodel4_pred, actual = listings3$loan_status)
logisticmodel4_conf_table
```

#### 4. Summary of Comparison Results

Comparison         |Model 2                         |Model 3                         |Model 4
-------------------|--------------------------------|--------------------------------|--------------------------------
AIC                |`r format(AIC_02, scientific=F)`|`r format(AIC_03, scientific=F)`|`r format(AIC_04, scientific=F)`
Pseudo R-Squared   |`r pR2_02`                      |`r pR2_03`                      |`r pR2_04`       
Prediction Accuracy|`r accuracy_02`                 |`r accuracy_03`                 |`r accuracy_04`  

### Model Interpretation

> ***Overall Fit***

* **Model 3** has relatively higher Pseudo R-squared among all candidates, so we expect this model to better predict further possibility of loan default (`loan_status`).
* The Pseudo R-squared of **Model 3** is `r pR2_03`, indicting only **`r paste(round(pR2_03*100, 2), "%", sep="")`** of the variation in the loan default probability (`loan_status`) in the sample is explained by the variation in a combination of loan term and size and credit riskvariables.
* Moreover, given the confusion matrix, when **cutoff = 0.1** (classify a borrower will default when the predicted probability is greater than or equal to 10% AND a borrower will not default when the predicted probability is less than or 10%), the overall accuracy of the prediction based on **Model 3** is **`r paste(round(accuracy_03*100, 2), "%", sep="")`**, suggesting the model does well in estimating the probability of default.

> ***Regression Coefficients***

```{r echo=FALSE, include=FALSE}
# All coefficients of Logistic Regression Model 3
logisticmodel3$coefficients

# Odds ratios for all independent variables of Logistic Regression Model 3
exp(coef(logisticmodel3))
```

##### A. Loan Size and Term

1. `number_of_days`: **odds ratio** = exp(coefficient of `number_of_days`) = `r round(exp(coef(summary(logisticmodel3))["number_of_days","Estimate"]), 4)`
* `number_of_days` **decreases** the probability of `dafault`: odds ratio of `number_of_days` is `r round(exp(coef(summary(logisticmodel3))["number_of_days","Estimate"]), 4)`, suggesting on average, controlling for all other variables, every one unit (one day) increase in the number of days of the loan **multiplies the odds of loan default** by `r round(exp(coef(summary(logisticmodel3))["number_of_days","Estimate"]), 4)`.

2. `principal_balance`: **odds ratio** = exp(coefficient of `principal_balance`) = `r round(exp(coef(summary(logisticmodel3))["principal_balance","Estimate"]), 6)`
* `principal_balance` **decreases** the probability of `dafault`: odds ratio of `principal_balance` is `r round(exp(coef(summary(logisticmodel3))["principal_balance","Estimate"]), 6)`, suggesting on average, controlling for all other variables, every one unit ($1) increase in the loan principal balance **multiplies the odds of loan default** by `r round(exp(coef(summary(logisticmodel3))["principal_balance","Estimate"]), 6)`.

3. `amount_funded`: **odds ratio** = exp(coefficient of `amount_funded`) = `r round(exp(coef(summary(logisticmodel3))["amount_funded","Estimate"]), 6)`
* `amount_funded` **increases** the probability of `dafault`: odds ratio of `amount_funded` is `r round(exp(coef(summary(logisticmodel3))["amount_funded","Estimate"]), 6)`, suggesting on average, controlling for all other variables, every one unit ($1) increase in the funded amount of the loan **multiplies the odds of loan default** by `r round(exp(coef(summary(logisticmodel3))["amount_funded","Estimate"]), 6)`.

4. `borrower_rate`: **odds ratio** = exp(coefficient of `borrower_rate`) = `r round(exp(coef(summary(logisticmodel3))["borrower_rate","Estimate"]), 4)`
* `borrower_rate` **increases** the probability of `dafault`: odds ratio of `borrower_rate` is `r round(exp(coef(summary(logisticmodel3))["borrower_rate","Estimate"]), 4)`, suggesting on average, controlling for all other variables, every one unit (1%) increase in the loan borrower rate **multiplies the odds of loan default** by `r round(exp(coef(summary(logisticmodel3))["borrower_rate","Estimate"]), 4)`.

5. `listing_term`: **odds ratio** = exp(coefficient of `listing_term`) = `r round(exp(coef(summary(logisticmodel3))["listing_term","Estimate"]), 4)`
* `listing_term` **increases** the probability of `dafault`: odds ratio of `listing_term` is `r round(exp(coef(summary(logisticmodel3))["listing_term","Estimate"]), 4)`, suggesting on average, controlling for all other variables, every one unit (one month) increase in the loan listing term **multiplies the odds of loan default** by `r round(exp(coef(summary(logisticmodel3))["listing_term","Estimate"]), 4)`.

##### B. Credit Score Matrices

1. `prosper_score`: **odds ratio** = exp(coefficient of `prosper_score`) = `r round(exp(coef(summary(logisticmodel3))["prosper_score","Estimate"]), 4)`
* `prosper_score` **decreases** the probability of `dafault`: odds ratio of `prosper_score` is `r round(exp(coef(summary(logisticmodel3))["prosper_score","Estimate"]), 4)`, suggesting on average, controlling for all other variables, every one unit increase in the prosper score **multiplies the odds of loan default** by `r round(exp(coef(summary(logisticmodel3))["prosper_score","Estimate"]), 4)`.

##### C. Other Credit Risk Matrices

1. `total_inquiries`: **odds ratio** = exp(coefficient of `total_inquiries`) = `r round(exp(coef(summary(logisticmodel3))["total_inquiries","Estimate"]), 4)`
* `total_inquiries` **increases** the probability of `dafault`: odds ratio of `total_inquiries` is `r round(exp(coef(summary(logisticmodel3))["total_inquiries","Estimate"]), 4)`, suggesting on average, controlling for all other variables, every one unit increase in the number of total inquiries **multiplies the odds of loan default** by `r round(exp(coef(summary(logisticmodel3))["total_inquiries","Estimate"]), 4)`.

2. `was_delinquent_derog`: **odds ratio** = exp(coefficient of `was_delinquent_derog`) = `r round(exp(coef(summary(logisticmodel3))["was_delinquent_derog","Estimate"]), 4)`
* `was_delinquent_derog` **decreases** the probability of `dafault`: odds ratio of `was_delinquent_derog` is `r round(exp(coef(summary(logisticmodel3))["was_delinquent_derog","Estimate"]), 4)`, suggesting on average, controlling for all other variables, every one unit increase in the previous delinquencies **multiplies the odds of loan default** by `r round(exp(coef(summary(logisticmodel3))["was_delinquent_derog","Estimate"]), 4)`.

3. `total_open_revolving_accounts`: **odds ratio** = exp(coefficient of `total_open_revolving_accounts`) = `r round(exp(coef(summary(logisticmodel3))["total_open_revolving_accounts","Estimate"]), 4)`
* `total_open_revolving_accounts` **increases** the probability of `dafault`: odds ratio of `total_open_revolving_accounts` is `r round(exp(coef(summary(logisticmodel3))["total_open_revolving_accounts","Estimate"]), 4)`, suggesting on average, controlling for all other variables, every one unit increase in total open revolving accounts **multiplies the odds of loan default** by `r round(exp(coef(summary(logisticmodel3))["total_open_revolving_accounts","Estimate"]), 4)`.

4. `revolving_balance`: **odds ratio** = exp(coefficient of `revolving_balance`) = `r round(exp(coef(summary(logisticmodel3))["revolving_balance","Estimate"]), 4)`
* `revolving_balance` **decreases** the probability of `dafault`: odds ratio of `revolving_balance` is `r round(exp(coef(summary(logisticmodel3))["revolving_balance","Estimate"]), 4)`, suggesting on average, controlling for all other variables, every one unit ($1) increase in the revolving balance **multiplies the odds of loan default** by `r round(exp(coef(summary(logisticmodel3))["revolving_balance","Estimate"]), 4)`.

5. `real_estate_balance`: **odds ratio** = exp(coefficient of `real_estate_balance`) = `r round(exp(coef(summary(logisticmodel3))["real_estate_balance","Estimate"]), 6)`
* `real_estate_balance` **decreases** the probability of `dafault`: odds ratio of `real_estate_balance` is `r round(exp(coef(summary(logisticmodel3))["real_estate_balance","Estimate"]), 6)`, suggesting on average, controlling for all other variables, every one unit ($1) increase in the real estate balance **multiplies the odds of loan default** by `r round(exp(coef(summary(logisticmodel3))["real_estate_balance","Estimate"]), 6)`.

***

## Analysis




***


## Conclusion

